\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-09-15 abstract.tex
isorec -> fix

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a language design useful for extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, a seemingly object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses significant technical challenges that arise
from the underlying language being simultaneously functional,
dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a formalization of the language design as a constructive type theory with inductive types,
\item proofs of strong guarantees afforded by the type theory,
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

>>>>

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a language design useful for extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, a seemingly object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses significant technical challenges that arise
from the underlying language of a proof assistant being simultaneously
functional, dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a formalization of the language design as a constructive type theory with inductive types,
\item proofs of strong guarantees afforded by the type theory,
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

### 2022-09-15 challenge.tex
isorec -> fix

\subsection{Challenges}

>>>>

Supporting inheritance and family polymorphism in a proof assistant poses challenges
far beyond those encountered in an object-oriented setting~\cite{ncm2004,vc-calculus-2006,zm2017},
because the underlying programming language is simultaneously
functional, dependently typed, a logic, and an interactive tool.

\noindentparagraph{Extensible inductive types vs.\ exhaustive induction reasoning.}

Inductive types, generalizing algebraic data types found in
functional languages, are a central feature of any proof assistant being
used for mechanizing language metatheories.
They offer a means to define abstract syntax, typing rules, reduction
rules, and so on.
Unfortunately, inductive types are closed to extension.

A family-polymorphism mechanism could potentially support extensible
inductive types, by allowing a \emph{derived family} to add new constructors
to inductive types inherited from a \emph{base family}.
Such capability would be useful for extending mechanized languages.
For example, \cref{fig:?} shows the part of an STLC extension that could
potentially be mechanized using such extensible inductive types: a derived family
inherits constructors of inductive types from the base family (viz.\ the $\cdots$'s)
and adds new constructors to model the new language feature.

\input{stlc-nonmechanized.tex}

However, there is a tension between extensibility of inductive types
and exhaustivity of induction reasoning.
In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
the substitution function, require induction reasoning.
In the derived family, to ensure that induction remains exhaustive,
how could the programmer account for the new cases arising from the new
constructors, without redefining or forcing the rechecking of those cases
already handled by the base family?

The tension results from the duality between variants and records.
For record-like language constructs (e.g., classes and objects), it is
safe for an extension to add more fields.
By contrast, variant-like language constructs (e.g., inductive types)
do not automatically enjoy safe, modular addition of constructors.

\noindentparagraph{Late binding vs.\ dependent typing.}

\noindentparagraph{Essence of inheritance vs.\ consistency of a logic.}

\noindentparagraph{Family vs.\ interactivity.}

\newpage
